---
title: "test final"
author: "Ben"
date: "9/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggmap)
library(tigris)
library(rnaturalearth)
library(tidycensus)
library(rgdal)
library(sf)
library(tmap)
library(lubridate)
library(vegan)
```

# Thinking about what I need to do
So with my iNaturalist data I think I need to apply the GEOID or some sort of identifier for the census block it is in but maybe keep it in a non-spatial format. I need that information and also likely statistics about each block group such as area. Then i can compute the biodiversity metrics for each block group and apply it to the polygons. 

So the only information I need from the block groups is the ID. I should apply that classification to each polygon that falls within it 




```{r}
d <- read_csv("final_project/data/in_929.csv",guess_max = 20000)
sapply(d,class)
```
```{r}
d <- d %>% select(id,observed_on, created_at, quality_grade,captive_cultivated,latitude, longitude,positional_accuracy,species_guess, scientific_name, common_name, iconic_taxon_name,taxon_genus_name)
```

```{r}
#census_api_key(key = "75eb1a4035ff5768216984592e3501956cdff75e", install = TRUE)
```

So I think I want to group inaturalist ovservations by census block group. Need to 

```{r}


block <- block_groups("massachusetts", "worcester")
```
```{r turn the INaturalist data into an sf object}
#converts it from df to sf and df
ds <- st_as_sf(d,coords = c(x = "longitude", y = "latitude"))

#assing the block group Coordinate refrence system to the inaturalist data 
st_crs(ds) <- st_crs(block)
```
```{r}
head(ds)
```

```{r}
st_area(block)
```
```{r distribution of Inaturalist observation over time}
ds %>%
  ggplot(aes(observed_on))+
  geom_freqpoly(binwidth = 31)
```
```{r add month and year column to ds}
ds <- ds %>% mutate(month = month(observed_on), year = year(observed_on))%>%dplyr::select(id, month, year, observed_on, everything())
```

should I use year specific observations or just total observations??? 
Year speciic would be good if i think things have changed much in the last three years 

```{r breakdown of obs per year}
ds %>% group_by(year) %>% summarize(obs = length(id)) %>% arrange(desc(obs))
```
```{r breakdown by taxon type }
ds %>% group_by(iconic_taxon_name) %>% summarize(obs = length(id)) %>% arrange(desc(obs))
```

```{r breakdown by taxon type }
ds %>% group_by(scientific_name) %>% summarize(obs = length(id)) %>% arrange(desc(obs))
```


```{r attempt at joining IN data with block group, then finding total number of obs and mapping it}
  
p <- st_join(block, ds) %>% group_by(GEOID) %>% summarize(n = length(id))%>% arrange(desc(n))

# filter out those w
not_enough_obs <- st_join(block, ds) %>% group_by(GEOID) %>% summarize(n = length(id))%>% arrange((n)) %>% filter(n  < 30 ) 

ggplot(p)+
  geom_sf(aes(fill = n)) +  
  geom_sf(data = not_enough_obs, aes())+
  coord_sf(xlim = c(-71.9, -71.7), ylim = c(42.2,42.35), expand = FALSE)
```

```{r attempt at joining IN data with block group, then finding total number of obs and mapping it}
  
obs_density <- st_join(block, ds) %>% group_by(GEOID, scientific_name) %>% mutate(n = length(scientific_name))%>% mutate(obs_dens = n/ALAND)

#try to make blocks with less than 50 obs empty 

ggplot(obs_density)+
  geom_sf(aes(fill = obs_dens)) +  
  coord_sf(xlim = c(-71.9, -71.7), ylim = c(42.2,42.35), expand = FALSE)
```

```{r find which block groups have enough iNaturalist obs}

geoid_enough_samples <- st_join(block, ds) %>% group_by(GEOID) %>% summarize(n = length(id))%>% arrange((n)) %>% filter(n  > 30 ) %>% pull(GEOID)

filtered_blocks_iN <- st_join(block, ds) %>% group_by(GEOID) %>% dplyr::filter(GEOID %in% geoid_enough_samples)
```

```{r richness of all inaturalist observations }
a <- filtered_blocks_iN %>% group_by(scientific_name) %>%
  mutate(n = length(scientific_name))%>%
  mutate(richness_density = n/ALAND)


b <- filtered_blocks_iN %>%
  group_by(GEOID) %>%
  mutate(richness = length(scientific_name)) %>%
  mutate(richness_density = (richness/ALAND)*1000) %>% dplyr::select(GEOID,BLKGRPCE,richness,richness_density, everything())

 ggplot()+
  geom_sf(data = b, aes(fill = richness_density)) +  
  geom_sf(data = not_enough_obs, aes()) +
  coord_sf(xlim = c(-71.9, -71.7), ylim = c(42.2,42.35), expand = FALSE)+  scale_fill_viridis_c()
```


```{r}
library(mapview)

mapView(b, zcol = "richness_density") 

```


```{r}
 test_join%>%  
  ggplot()+
  geom_sf() + 
  geom_sf(ds)+
  coord_sf(xlim = c(-71.9, -71.7), ylim = c(42.2,42.35), expand = FALSE)

```
```{r}
tm_shape(test_join)+
  tm_polygons()
```


```{r}

  ggplot()+
  geom_sf(data = test_join) + 
  geom_sf(data = ds, aes(color = iconic_taxon_name),alpha = .5)+
    coord_sf(xlim = c(-71.9, -71.7), ylim = c(42.2,42.35), expand = FALSE)
```




