---
title: "w3_analysis"
author: "Ben"
date: "9/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(janitor)
library(ggrepel)
library(gt)
library(grid)
library(broom)

```

```{r}
urban <-read_csv("./w3/data/urban_data.csv")
```
```{r}
#clean names
urban <-  clean_names(urban)
```

```{r find values needing cleaning}
#find unique values to see if value coding shoudl be corrected
sapply(urban, unique)
#looks like only the site neds to be fixed
```

```{r clean site names}
#clean site names
urban_tidy <- urban %>%
  mutate(site = case_when(site == "LOT" ~ "lot",
                              site == "Lot" ~ "lot",
                              site == "Park" ~ "park",
                              site == "PARK" ~ "park"))

```

```{r create two seperate data sets for lm}

lot <- urban_tidy %>% filter(site == "lot")

park <- urban_tidy %>% filter(site == "park")
```

```{r get names of numeric variables for lm}
v_names <- names(select_if(urban_tidy, is.numeric))
```

```{r}
#models <- lapply(v_names, function(x) {
   # lm(substitute(read ~ i, list(i = as.name(x))), data = hsb2)
#})
```

varlist <- names(hsb2)[8:11]

models <- lapply(varlist, function(x) {
    lm(substitute(read ~ i, list(i = as.name(x))), data = hsb2)
})

## look at the first element of the list, model 1
models[[1]]
```{r example code i try to copy }
v_names
```
```{r}
#lot and flora
lm(data = lot, lot$flora_number_spp ~ 
     debris_number +
     infiltration_sec_values_60_are_actualy_60 +
     o_horizon_cm_values_1_are_actually_below_1 + 
     cover_class_1_percent + cover_class_2_percent +
     cover_class_3_percent ) %>%
  summary
  
```
```{r}
#park and flora
lm(data = park, flora_number_spp ~ 
     debris_number +
     infiltration_sec_values_60_are_actualy_60 +
     o_horizon_cm_values_1_are_actually_below_1 + 
     cover_class_1_percent + cover_class_2_percent +
     cover_class_3_percent ) %>%
  summary()

```

```{r}
#park and fauna species
lm(data = park,  park$fauna_number_spp~ 
     debris_number +
     infiltration_sec_values_60_are_actualy_60 +
     o_horizon_cm_values_1_are_actually_below_1 + 
     cover_class_1_percent + cover_class_2_percent +
     cover_class_3_percent ) %>%
  summary()
```

```{r}
#park and fauna individuals
lm(data = park,  park$fauna_number_individuals_values_25_are_actually_above_25~ 
     debris_number +
     infiltration_sec_values_60_are_actualy_60 +
     o_horizon_cm_values_1_are_actually_below_1 + 
     cover_class_1_percent + cover_class_2_percent +
     cover_class_3_percent ) %>%
  summary()

```

```{r}
#lot and fauna species
lm(data = lot,  lot$fauna_number_spp~ 
     debris_number +
     infiltration_sec_values_60_are_actualy_60 +
     o_horizon_cm_values_1_are_actually_below_1 + 
     cover_class_1_percent + cover_class_2_percent +
     cover_class_3_percent ) %>%
  summary()
```

```{r lms }

#lot and fauna individuals
lm(data = lot,  lot$fauna_number_individuals_values_25_are_actually_above_25~ 
     debris_number +
     infiltration_sec_values_60_are_actualy_60 +
     o_horizon_cm_values_1_are_actually_below_1 + 
     cover_class_1_percent + cover_class_2_percent +
     cover_class_3_percent ) %>%
  summary()


```

1.	Include figures to link each 


- independent variable 
    (infiltration, 
    litter, 
    depth of soil O-horizon (cm), 
    soil color (based on Munsell chart), 
    heterogeneity of cover classes) 
  to each 
  dependent variable 
    (number of unique species of flora, 
    number of individuals and 
    unique species of fauna).

Discussion – include discussion of the following questions: 

1.	How do floral and faunal communities differ between the two study sites?
2.	What variables best explain floral diversity?
3.	What variables best describe faunal abundance and diversity?
4.	What limitations are there of our research design?  How could it be improved?

# matching independant and dependant variables
## flora
flora + infiltration
flora + litter
flora + soil depth
flora + soil color
flora + heterogeneity of cover classes


```{r flora + infiltration}


urban_tidy %>%
  ggplot(aes(y =flora_number_spp,
             x = infiltration_sec_values_60_are_actualy_60, 
             color = site))+
  geom_point( )+
  geom_smooth(method = "lm")+
  labs(x = "Infiltration (sec) [Values = 60 actually exceeded 60] ", y = "Flora number of unique species")
ggsave("w3/flora_infiltration.png")


```

flora + litter

```{r flotra + litter}
urban_tidy %>%
  ggplot(aes(y =flora_number_spp,
  x = debris_number, color = site))+
  geom_point( ) + 
  geom_smooth(method = "lm")+
  labs(x = "# of debris", y = "Flora number of unique species")
ggsave("w3/flora_debris.png")

```
flora + soil depth
```{r flora + soil depth}
urban_tidy %>%
  ggplot(aes(y =flora_number_spp,
  x = o_horizon_cm_values_1_are_actually_below_1, 
             color = site))+
  geom_point( )+
  geom_smooth(method = "lm")+
  labs(x = "depth of soil O-horizon (cm) [values = 1, are actually below 1] ", y = "Flora number of unique species")
ggsave("w3/flora_soildepth.png")
urban_tidy
```

flora + soil color
```{r flora + soil color}
urban_tidy %>%
  ggplot(aes(y =flora_number_spp,
    x = soil_color_munsell, 
             fill = site))+
  geom_bar(sta ="identity")+ 
  coord_flip()+ 
  labs(x = "Soil color (munsell)", y= "# of unique flora species")
ggsave("w3/flora_sc.png")
```
flora + heterogeneity of cover classes - not sure how to figure this out. NEED TO DO SOME TRANSFORMATION
```{r flora + heterogeneity of cover classes}
urban_tidy %>%
  ggplot(aes(y =flora_number_spp,
  x = cover_class_1_percent, 
             color = site))+
  geom_point( )+
  geom_smooth(method = "lm")
```
```{r covr class 2flora }
urban_tidy %>%
  ggplot(aes(y =flora_number_spp,
  x = cover_class_2_percent, 
             color = site))+
  geom_point( )+
  geom_smooth(method = "lm", alpha = .2)+labs(x = "% Cover class 2", y = "# of unique species of flora")
ggsave("w3/floraclass2.png")


urban_tidy %>%
  ggplot(aes(y =flora_number_spp,
  x = cover_class_3_percent, 
             color = site))+
  geom_point( )+
  geom_smooth(method = "lm", alpha = .2)+labs(x = "% Cover class 3", y = "# of unique species of flora")
ggsave("w3/floraclass3.png")
```
```{r}

```

## fauna number
fauna + infiltration
fauna + litter
fauna + soil depth
fauna + soil color
fauna + heterogeneity of cover classes
```{r}


```


```{r fauna + infiltration}


urban_tidy %>%
  ggplot(aes(y =fauna_number_spp,x = infiltration_sec_values_60_are_actualy_60, 
             color = site))+
  geom_point( )+
  geom_smooth(method = "lm")+ 
  labs(x = "Infiltration (sec) [Values = 60 actually exceeded 60] ", y= "# of unique fauna species")
ggsave("w3/faunasp_infiltration.png")


```

fauna + litter

```{r fauna + litter}

urban_tidy %>%
  ggplot(aes(y =fauna_number_spp,
  x = debris_number, 
             color = site))+
  geom_point( )+
  geom_smooth(method = "lm")+ 
  labs(x = "debirs #", y= "# of unique fauna species")
ggsave("w3/faunasp_debris.png")

```
fauna species + soil depth
```{r fauna + soil depth}
urban_tidy %>%
  ggplot(aes(y =fauna_number_spp,
  x = o_horizon_cm_values_1_are_actually_below_1, 
             color = site))+
  geom_point( )+
  geom_smooth(method = "lm")+
  labs(x = "depth of soil O-horizon (cm)", y = "# of unique Fauna species")
ggsave("w3/faunasp_soildepth.png")
```
fauna + soil color

```{r fauna species and soil color }
urban_tidy %>%
  ggplot(aes(y =fauna_number_spp,
    x = soil_color_munsell, 
             fill = site))+
  geom_bar(sta ="identity")+ 
  coord_flip()+ 
  labs(x = "Soil color (munsell)", y= "# of unique fauna species")
ggsave("w3/fauna_sc.png")
```
fauna + heterogeneity of cover classes - not sure how to figure this out. NEED TO DO SOME TRANSFORMATION
```{r fauna diversity + heterogeneity of cover classes}
urban_tidy %>%
  ggplot(aes(y =fauna_number_spp,
  x = cover_class_1_percent, 
             color = site))+
  geom_point( )+
  geom_smooth(method = "lm", alpha = .2)+
  labs(x = "% Cover class 1", y = "# of unique species of fauna")
ggsave("w3/faunaclass1.png")

urban_tidy %>%
  ggplot(aes(y =fauna_number_spp,
  x = cover_class_2_percent, 
             color = site))+
  geom_point( )+
  geom_smooth(method = "lm", alpha = .2)+
  labs(x = "% Cover class 2", y = "# of unique species of fauna")
ggsave("w3/faunaclass2.png")

urban_tidy %>%
  ggplot(aes(y =fauna_number_spp,
  x = cover_class_3_percent, 
             color = site))+
  geom_point( )+
  geom_smooth(method = "lm")+
  labs(x = "% Cover class 3", y = "# of unique species of fauna")
ggsave("w3/faunaclass3.png")

```

```{r}
urban_tidy %>% group_by(soil_color_munsell)%>%
  summarise(count = n())
```

## number of individual of fauna
fauna infiltration
```{r fauna infiltration}

urban_tidy %>%
  ggplot(aes(y =fauna_number_individuals_values_25_are_actually_above_25,x = infiltration_sec_values_60_are_actualy_60, 
             color = site))+
  geom_point( )+
  geom_smooth(method = "lm")+ 
  labs(x = "Infiltration (sec) [Values = 60 actually exceeded 60] ", y= "# of fauna individuals")
ggsave("w3/fauna_ind_infiltration.png")

```



fauna + soil depth
```{r fauna + soil depth}
urban_tidy %>%
  ggplot(aes(y =fauna_number_individuals_values_25_are_actually_above_25,
  x = o_horizon_cm_values_1_are_actually_below_1, 
             color = site))+
  geom_point( )+
  geom_smooth(method = "lm", alpha = .3)+
  labs(x = "depth of soil O-horizon (cm)", y = "# of Fauna individuals")
ggsave("w3/fi_soil.png")

```


```{r fauna + litter}

urban_tidy %>%
  ggplot(aes(y =fauna_number_individuals_values_25_are_actually_above_25,
  x = debris_number, 
             color = site))+
  geom_point( )+
  geom_smooth(method = "lm")+ 
  labs(x = "debris #", y= "# of Fauna individuals")
ggsave("w3/fi_debris.png")

```

```{r fauna individuals and soil color }
urban_tidy %>%
  ggplot(aes(y =fauna_number_individuals_values_25_are_actually_above_25,
    x = soil_color_munsell, 
             fill = site))+
  geom_bar(sta ="identity")+ 
  coord_flip()+ 
  labs(x = "Soil color (munsell)", y= "# of fauna individuals")
ggsave("w3/fi_sc.png")
```

```{r fauna diversity + heterogeneity of cover classes}
urban_tidy %>%
  ggplot(aes(y =fauna_number_individuals_values_25_are_actually_above_25,
  x = cover_class_1_percent, 
             color = site))+
  geom_point( )+
  geom_smooth(method = "lm", alpha = .2)+
  labs(x = "% Cover class 1", y = "# of fauna individuals")
ggsave("w3/ficlass1.png")

urban_tidy %>%
  ggplot(aes(y =fauna_number_individuals_values_25_are_actually_above_25,
  x = cover_class_2_percent, 
             color = site))+
  geom_point( )+
  geom_smooth(method = "lm", alpha = .2)+
  labs(x = "% Cover class 2", y = "# of fauna individuals")
ggsave("w3/ficlass2.png")

urban_tidy %>%
  ggplot(aes(y =fauna_number_individuals_values_25_are_actually_above_25,
  x = cover_class_3_percent, 
             color = site))+
  geom_point( )+
  geom_smooth(method = "lm")+
  labs(x = "% Cover class 3", y = "# of fauna individuals")
ggsave("w3/ficlass3.png")

```

```{r}
summary(park)
```
```{r}
summary(lot)
```
```{r}
n <- urban_tidy %>% group_by(site) %>%names()
  ggplot(data = urban_tidy)+
  boxplot(aes(x = n))
```

