---
title: "w3_analysis"
author: "Ben"
date: "9/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(janitor)
library(ggrepel)
library(gt)
library(grid)

```

```{r}
urban <-read_csv("./w3/data/urban_data.csv")
```
```{r}
#clean names
urban <-  clean_names(urban)
```

```{r find values needing cleaning}
#find unique values to see if value coding shoudl be corrected
sapply(urban, unique)
#looks like only the site neds to be fixed
```

```{r clean site names}
#clean site names
urban_tidy <- urban %>%
  mutate(site = case_when(site == "LOT" ~ "lot",
                              site == "Lot" ~ "lot",
                              site == "Park" ~ "park",
                              site == "PARK" ~ "park"))

```
1.	Include figures to link each 


- independent variable 
    (infiltration, 
    litter, 
    depth of soil O-horizon (cm), 
    soil color (based on Munsell chart), 
    heterogeneity of cover classes) 
  to each 
  dependent variable 
    (number of unique species of flora, 
    number of individuals and 
    unique species of fauna).

Discussion – include discussion of the following questions: 

1.	How do floral and faunal communities differ between the two study sites?
2.	What variables best explain floral diversity?
3.	What variables best describe faunal abundance and diversity?
4.	What limitations are there of our research design?  How could it be improved?


```{r}
names(urban_tidy)

urban_tidy %>%
  ggplot(aes(y =flora_number_spp))+
  geom_point(x = infiltration_sec_values_60_are_actualy_60 )



```

